---
title: "PML Project"
author: "J ten Veen"
date: "4-3-2018"
output: html_document
---

## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Load packages
```{r}
##Load packages
library(caret)
library(rpart.plot)
library(randomForest)
```

## Data processing
#Load data when present in working directory.
```{r}
training_pml <- read.csv("pml-training.csv", na.strings = c("NA", ""))
testing_pml <- read.csv("pml-testing.csv", na.strings = c("NA", ""))
```

Show structure of the data.
```{r}
dim(training_pml)
```
A data frame with 19622 observations of 160 variables.

## Cleaning the data
Remove columns with NA missing values.
```{r}
training_pml <- training_pml[, colSums(is.na(training_pml)) == 0]
testing_pml <- testing_pml[, colSums(is.na(testing_pml)) == 0] 
```

Remove identification only variables (columns 1 to 5).
```{r}
identification <- names(training_pml) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp")
training_pml <- training_pml[!identification]
```

## Data splitting
The test dataset does not contain the classe variable. These classes should be predicted by the model. Therefore we split the training data within 2 groups. Training (train_data) and testing (test_data).
```{r}
set.seed(50)
inTrain <- createDataPartition(training_pml$classe, p=0.70, list=F)
train_data <- training_pml[inTrain, ]
test_data <- training_pml[-inTrain, ]
dim(train_data)
dim(test_data)
```

## Model building
### Decision Tree
```{r}
set.seed(50)
modelfit_dt <- rpart(classe ~ ., data=train_data, method="class")
prp(modelfit_dt)
```

Predict on test_data
```{r}
predict_dt <- predict(modelfit_dt, newdata=test_data, type="class", trControl = control)
results_dt <- confusionMatrix(predict_dt, test_data$classe)
results_dt
```
It is shown that the Accuracy is .7594. The expected out of sample error is 0.2406. Therefore, we'll try another method.

## Use random forest
```{r}
modelfit_rf <- randomForest(classe ~., data=train_data)
result_rf <- confusionMatrix(test_data$classe, predict(modelfit_rf, newdata=test_data))
result_rf
```
It is shown that the Accuracy is .9963, which is better than the decision tree model. Therefore, the expected out of sample error rate is 0.0037.

## Conclusion
Based on the two models. We can conclude that, in this case, the Random Forest Classification works better than the Decision Tree Method. The results obtained by the Random Forest Classification were highly accurate on the used dataset. A disadvantage of this technique is the loading time. However, we used the RF model to predict the 20 cases.